---
title: "Project 4"
output: html_document
date: "2022-12-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## System I 


```{r}
myurl = "https://liangfgithub.github.io/MovieData/"
ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)

colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
library(dplyr)


```

```{r}
movies = readLines(paste0(myurl, 'movies.dat?raw=true'))
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)

# convert accented characters
movies$Title[73]
movies$Title = iconv(movies$Title, "latin1", "UTF-8")
movies$Title[73]

# extract year
movies$Year = as.numeric(unlist(
  lapply(movies$Title, function(x) substr(x, nchar(x)-4, nchar(x)-1))))
```


```{r}
tmp = ratings %>% 
  group_by(MovieID) %>% 
  summarize(ratings_per_movie = n(), ave_ratings = mean(Rating)) %>%
  inner_join(movies, by = 'MovieID')
summary(tmp$ratings_per_movie)
```


```{r}
pop_recom <- tmp %>% 
  filter(ratings_per_movie > 50) %>%
  arrange(desc = ratings_per_movie) %>%
  group_by(Genres) %>%
  slice(1:50) %>%
  select(c("Title", "Genres", "ratings_per_movie")) 
```

```{r}
library(ggplot2)
pop_recom  %>% ggplot(aes(ratings_per_movie)) + 
  geom_bar(fill = "steelblue", width = 1) + coord_cartesian(c(1,1500))
```
```{r}
pop_recom %>% group_by(Genres) %>% summarise(movie_count = n())
```



The Popularity based recommendations are built upon the movies that have at least 50 reviews and are ordered from highest score to the lowest, clipping at index of 50. In other words, the recommendation guarantees in each Genre, the model produces at most 50 movies that received at least 50 reviews. 

I chose to do this as popular movies are, by definition, have obtained some exposure to the public and have been rated highly.

### Recommending the most trendy movies

```{r}
tmp$weight_fact = tmp$Year * tmp$ave_ratings


tmp %>% 
  filter(ratings_per_movie > 50) %>%
  arrange(desc(Year), desc(weight_fact)) %>%
  group_by(Genres) %>%
  slice(1:50) %>%
  select(c("Title", "Genres", "ratings_per_movie", "Year")) 

```

To recommend the most trendy movies, I used a weight factor that takes the average rating and year into account. The model should output the highest ranking of movie of the most recent years. Similar to previous method, the movies requires at least 50 ratings are received. 




## System II - Collaborative Filtering

### User Based CF
```{r}
library(recommenderlab)
myurl = "https://liangfgithub.github.io/MovieData/"
ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
i = paste0('u', ratings$UserID)
j = paste0('m', ratings$MovieID)
x = ratings$Rating
tmp = data.frame(i, j, x, stringsAsFactors = T)
Rmat = sparseMatrix(as.integer(tmp$i), as.integer(tmp$j), x = tmp$x)
rownames(Rmat) = levels(tmp$i)
colnames(Rmat) = levels(tmp$j)
Rmat = new('realRatingMatrix', data = Rmat)

```


```{r}
library(proxy)
train = Rmat[1:500, ]
test = Rmat[501, ]
data = as(train, "matrix")
user.means = rowMeans(data, na.rm = TRUE)
data = data - user.means
newdata = as(Rmat[501, ], "matrix")
newuser.mean = mean(newdata, na.rm = TRUE)
newdata = newdata - newuser.mean
sim1 = proxy::simil(data, newdata, method = "cosine")
sim1 = (1 + sim1)/2
sim1 = sim1[, 1]

neigh = tail(order(sim1, decreasing = FALSE, na.last=NA), 20)
tmp_data = data[neigh, ]

numerator   = colSums(tmp_data*sim1[neigh], na.rm=TRUE) 
denominator = colSums((!is.na(tmp_data))*sim1[neigh])

mypred = numerator/denominator 

mypred = mypred + newuser.mean
mypred[is.nan(mypred)] <- NA    # abnormal results
mypred[is.infinite(mypred)] <- NA   # abnormal results
mypred[!is.na(newdata)]<-NA   # already watched movies

recommender.UBCF <- Recommender(train, method = "UBCF",
                                parameter = list(normalize = 'center', 
                                                 method = 'Cosine', 
                                                 nn = 20))

p.UBCF <- predict(recommender.UBCF, test, type="ratings")
p.UBCF <- as.numeric(as(p.UBCF, "matrix"))

sum(is.na(p.UBCF) != is.na(mypred)) ### should be zero
max(abs(p.UBCF - mypred), na.rm = TRUE)  ### should be less than 1e-06 
```



### Item Based CF

```{r}

train = Rmat[1:500, ]
data = as(train, "matrix")
user.means = rowMeans(data, na.rm = TRUE)
data = data - user.means

test = Rmat[501, ]
newdata = as(test, "matrix")

```

```{r}
item_sim = proxy::simil(data, by_rows = FALSE, method = "cosine")
item_sim = (1 + item_sim)/2
item_sim = as.matrix(item_sim)
```


```{r}
item.sim.top30 = matrix(0, nrow = dim(data)[2], ncol = dim(data)[2])
# j = paste0('m', ratings$MovieID)
rownames(item.sim.top30) = colnames(data)
colnames(item.sim.top30) = colnames(data)
```


```{r zero_fill}
for (i in 1:dim(item_sim)[1]){
  neighbor_ids = tail(order(item_sim[i,], decreasing = FALSE, na.last=NA), 30)
  item.sim.top30[i, neighbor_ids] = item_sim[i,neighbor_ids]
  antiset = setdiff( 1:dim(data)[2], neighbor_ids)
  item.sim.top30[i, antiset] = 0
  item.sim.top30[i,i] = 0
}
```


```{r gt_alternative, include=FALSE, echo=FALSE, eval=FALSE}
# u <- normalize(newdata, method = recommender.IBCF@model$normalize)
sim2 <- recommender.IBCF@model$sim
u <- as(test, "dgCMatrix")

## predict all ratings (average similarity)
#ratings <- tcrossprod(sim,u)
ratings <-
  #t(as(tcrossprod(sim, u) / tcrossprod(sim != 0, u != 0), "matrix"))
  t(as(tcrossprod(sim2, u), "matrix") / as(tcrossprod(sim2 != 0, u != 0), "matrix"))
dimnames(ratings) <- dimnames(newdata)
mypred <- as.numeric(as(ratings, "matrix"))
mypred[is.nan(mypred)] <- NA    # abnormal results
mypred[is.infinite(mypred)] <- NA   # abnormal results
mypred[!is.na(newdata)]<-NA   # already watched movies
```

```{r best}
newdata.nonNA = which(!is.na(newdata))
sim3 = item.sim.top30[,newdata.nonNA] 
u = t(as.matrix(newdata[newdata.nonNA]))
x = tcrossprod(sim3, u)
y = 1/rowSums(sim3, na.rm = TRUE)
mypreds2 = t(x * y)
mypreds2 = as.numeric(as(mypreds2, "matrix"))

mypreds2[is.nan(mypreds2)] <- NA    # abnormal results
mypreds2[is.infinite(mypreds2)] <- NA   # abnormal results
mypreds2[!is.na(newdata)]<-NA   # already watched movies
mypred = mypreds2

```

```{r eval_code1}
recommender.IBCF <- Recommender(train, method = "IBCF",
                                parameter = list(normalize = 'center', 
                                                 method = 'Cosine', 
                                                 k = 30))

p.IBCF <- predict(recommender.IBCF, test, type="ratings")
p.IBCF <- as.numeric(as(p.IBCF, "matrix"))

```

```{r eval_code2}

## first output: should be less than 10
sum(is.na(p.IBCF) != is.na(mypred))  

## second output: should be less than 10%
mydiff = abs(p.IBCF - mypred)
sum(mydiff[!is.na(mydiff)] > 1e-6) / sum(!is.na(mydiff)) 
```

